---
title: "How one year of Trump's 'Liberation Day' tariffs has changed global trade"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
---

# Introduction

One year after the Trump’s "Liberation Day" tariffs, global trade flows have shifted. Trade data shows which countries gained, which lost – and who is footing the bill.

*In this repository, you will find the methodology, data and code behind the stories that came out of this analysis.*

**Read the full story here:** [English](https://p.dw.com/p/xxx) \| [German](https://p.dw.com/p/xxx)

**Story by:** [Kira Schacht](https://www.dw.com/en/kira-schacht/person-46893544)

# Overview

We'll look at US import data provided through [ITC Trade Map](https://www.trademap.org/) to analyze:

-   How overall imports to the US have changed
-   Which countries have seen additional imports in various phases of the implementation, e.g.: -- January to March 2025: Stockpiling in anticipation of tariffs -- April to July 2025: Import substitution during between announcement (April 2) and implementation (August 7)
-   Customs duties collected by the US treasury over time, collated by the [Penn Wharton Budget Model](https://budgetmodel.wharton.upenn.edu/data-interactives/real-time-federal-budget-tracker.md/?_inputs_&budget_category=%22Taxes%22&budget_category_mirror=%22Taxes%22&maintabs=%22Tracker%22&budget_item=%22Customs%20Duties%20and%20Related%20Taxes%22&years=%2210%22&cumulative=%22TRUE%22&daily_freq=%22TRUE%22&inflation_adjust=%22TRUE%22&comparison_value=%22Previous%20Year%22&chart_tooltip_show_all=false)

# Analysis

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = '/Users/SchachtK/Library/CloudStorage/OneDrive-DeutscheWelle/_Projects/2601 trump tariffs/extern/data') #INSERT YOUR PROJECT DIRECTORY HERE
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
```

## Read Data: Monthly US imports

ITC Trade Map provides the monthly import volume to the US by country.

We'll focus on the years 2022-2024 as a baseline for comparison with 2025, since 2021 imports might still have been skewed by the Covid-19 pandemic.

```{r clean-trade-data, echo=FALSE}
#clean us trade data

#join 3 exported tables from Trade Map together
lf = list.files("raw", "Trade_Map_-_List_of_supplying_markets_for_a_product_imported_by_United_States_of_America_2", full.names = T)
d1 = read.delim(lf[1]); d2 = read.delim(lf[2]); d3 = read.delim(lf[3])
d = full_join(d1, d2) %>% full_join(d3)
rm(lf, d1, d2, d3)

#clean dataset
d = d %>%
  #pivot table to move months to rows instead of columns
  pivot_longer(2:last_col()) %>% 
  #remove empty rows
  filter(name != "X") %>% 
  #clean month values, separate into year and month category, convert unit to single dollars
  mutate(date = name %>% gsub("Imported.value.in.","",.) %>% ym,
         year = date %>% year(),
         month = date %>% month(),
         value = value * 1000) %>% 
  mutate(Exporters = Exporters %>% #fix country names
           gsub("Taipei, Chinese", "Taiwan", .) %>% 
           gsub("Korea, Republic of", "South Korea", .) %>% 
           gsub("Viet Nam", "Vietnam", .)) %>% 
  filter(year >= 2022) %>% #omit 2021 months
  select(-name, -date) #remove original date columns


#historic import shares 2022-2024
d_shares_historic = d %>%
  filter(Exporters != "World", year < 2025) %>% 
  group_by(Exporters) %>% 
  summarise(value = sum(value)) %>% 
  mutate(share = value / sum(value),
         g1p = share > 0.01)
#list of countries with historic shares over 1 percent
countries_g1p = d_shares_historic$Exporters[d_shares_historic$g1p]

save.image("data.RData")

d %>% 
  mutate(value = value/10^9) %>% head(n = 12) %>%
  knitr::kable(digits = 2, caption = "data preview, imports to the US in USD bn:")
```

We'll focus our analysis on the **`r length(countries_g1p)`** countries that trade the most with the US (make up at least 1% of US imports between 2022 and 2024):

**`r paste(countries_g1p, collapse = ", ")`**

## Total US import deviations

We'll compare total monthly US import values for 2025 to the average in the corresponding months of 2022-2024. The results show that total import values to the US briefly rose in early 2025, then returned to normal. We'll export this result to `trademap_us_imports_deviation.csv`.

```{r imports-deviation}
# us imports 2025 deviation from average
d1_total_deviation = d %>%
  filter(Exporters == "World") %>% 
  select(-Exporters) %>% 
  mutate(tmp = ifelse(year == 2025, "2025", "average")) %>% 
  group_by(tmp, month) %>% 
  summarise(value = mean(value)) %>% 
  pivot_wider(names_from = tmp, values_from = value) %>% 
  mutate(deviation = `2025`-average,
         month = month(month, label = T))

write.csv(d1_total_deviation, "processed/trademap_us_imports_deviation.csv",na = "", row.names = F)

d1_total_deviation %>% 
  mutate_at(2:4, ~./10^9) %>% 
  knitr::kable(digits = 2, caption = "Q1 additional imports, in USD bn")
```

## Q1 2025: Import Surges by country

In early 2025, anticipation of tariffs led US importers to stockpile goods from many trading partners. Focusing on the countries with more than 1% average import share to the US, we'll show the magnitude of additional imports in Q1 2025 compared to the average of Q1 2022-2024 per country.

The result is exported to `trademap_us_exporters_deviation_Q1.csv`

```{r q1-surges}
# countries with the biggest import surges:
# Q1 2025 vs average of Q1 2022-2024
d2_q1_deviation = d %>%
  filter(month <= 3, Exporters %in% countries_g1p) %>% 
  #sum Q1 of each year
  group_by(Exporters, year) %>% 
  summarise(value = sum(value)) %>% 
  #compare Q1 2025 with previous years
  mutate(tmp = ifelse(year == 2025, "Q1_2025", "Q1_average")) %>%
  group_by(Exporters, tmp) %>% 
  summarise(value = mean(value)) %>% 
  #restructure to move years to columns
  pivot_wider(names_from = tmp, values_from = value) %>% 
  #calculate absolute deviation and percentage change
  mutate(deviation = Q1_2025 - Q1_average,
         change_percent = Q1_2025 / Q1_average)

write.csv(d2_q1_deviation, "processed/trademap_us_exporters_deviation_Q1.csv",na = "", row.names = F)

d2_q1_deviation %>% 
  arrange(desc(deviation)) %>% 
  head(10) %>% 
  mutate_at(2:4, ~./10^9) %>% 
  knitr::kable(digits = 2, caption = "Q1 additional imports, in USD bn")
```

## April-July 2025: Import shifts during tariff pause

Between the announcement of tariffs on April 2 and their implementation in August, importers tried to shift to lower-tariff countries.

We'll show the resulting shift in imports on a map, the data of which is exported in `trademap_us_exporters_deviation_Apr_Jul.csv`

```{r apr-jul-surges}
# countries with the biggest import surges during tariff pause:
# Apr-Jul 2025 vs average of Apr-Jul 2022-2024
d3_apr_jul_deviation = d %>%
  filter(month %in% 4:7) %>% 
  group_by(Exporters, year) %>% #sum months of each year
  summarise(value = sum(value)) %>% 
  mutate(tmp = ifelse(year == 2025, "AJ_2025", "AJ_average")) %>% 
  group_by(Exporters, tmp) %>% 
  summarise(value = mean(value)) %>% 
  pivot_wider(names_from = tmp, values_from = value) %>% 
  mutate(deviation = AJ_2025 - AJ_average,
         change_percent = AJ_2025 / AJ_average)

write.csv(d3_apr_jul_deviation,  "processed/trademap_us_exporters_deviation_Apr_Jul.csv", na = "", row.names = F)

d3_apr_jul_deviation %>% 
  arrange(desc(deviation)) %>% 
  head(12) %>% 
  mutate_at(2:4, ~./10^9) %>% 
  knitr::kable(digits = 2, caption = "April-July change in imports, in USD bn")
```

## Monthly import changes in 2025

For a view of the year 2025 in total, we'll show the monthly deviation from the 2022-2024 average for all countries with at least 1% historic import share.

```{r monthly-changes}
# import change in total value during 2025 compared to 2022-2024 average
# countries with at least 1 percent market share
tmp = d %>% filter(Exporters %in% countries_g1p) %>% 
  mutate(tmp = ifelse(year == 2025, "2025", "average")) %>% 
  group_by(Exporters, month, tmp) %>% 
  summarise(value = mean(value)) %>% 
  pivot_wider(names_from = tmp, values_from = value) %>% 
  mutate(deviation = `2025` - average) %>% 
  select(Exporters, month, deviation) %>% 
  arrange(desc(deviation))

#get order of countries: biggest to smallest deviation april-july
tmp_countries = tmp %>%
  filter(month %in% 4:7) %>% group_by(Exporters) %>%
  summarise(deviation = sum(deviation)) %>% arrange(desc(deviation)) %>%
  select(Exporters) %>% `[[`(1)

d4_monthly_deviation = tmp %>% 
  mutate(Exporters = factor(Exporters, levels = tmp_countries),
         month = month(month, label = T)) %>% 
  arrange(Exporters) %>% 
  pivot_wider(names_from = Exporters, values_from = deviation)
rm(tmp,tmp_countries)

write.csv(d4_monthly_deviation, "processed/trademap_us_exporters_deviation_monthly.csv",na = "", row.names = F)

d4_monthly_deviation %>% 
  select(1:10) %>% 
  mutate_at(2:10, ~./10^9) %>% 
  knitr::kable(digits = 2, caption = "Preview: Monthly deviations by country, in USD bn")
```

## Penn Wharton Budget Model: Customs duties collected by US Treasury

`pennwharton_budget_model_taxes.csv` shows weekly amounts of customs duties collected by the US, processed from the data source for visualization.

```{r clean-penn-wharton}
# create a dataframe of week start dates for 2025
weekdates = read.csv("raw/Taxes_Customs Duties and Related Taxes_Weekly_2016-2026.csv") %>% 
  filter(Year == 2025) %>% 
  select(Week, Date)

d5_cutoms_duties = read.csv("raw/Taxes_Customs Duties and Related Taxes_Weekly_2016-2026.csv") %>% 
  select(Year, Week, Amount_YTD_Adj) %>% 
  #append dates to week numbers
  left_join(weekdates, by = join_by(Week)) %>%
  #clean data
  filter(!is.na(Date)) %>% arrange(Year, Week) %>% select(-Week) %>% 
  #one column per year
  pivot_wider(names_from = Year, values_from = Amount_YTD_Adj) %>% 
  #let the first week of 2025 begin in January instead of December
  mutate(Date = gsub("2025-12-31","2025-01-01",Date))

write.csv(d5_cutoms_duties, "processed/pennwharton_budget_model_taxes.csv",
          row.names = F, na = "")
rm(weekdates)

```
